---
title: "Chapter 1"
output: html_document
date: "2018-01-01"
---

###Financial Modeling 

#### Chapter 1

Calculating the present value of with equal payments

```{r pv equal pv, warning=FALSE,message=FALSE}
library(FinCal)
library(tidyquant)
library(jrvFinance)
library(ggthemes)
library(dygraphs)
library(timetk)
library(knitr)
library(kableExtra)


r <- 0.1
cf <- tribble(~Year,~cf,
              1,100,
              2,100,
              3,100,
              4,100,
              5,100)
cf %>%
  mutate(pv = cf/(1+r)^Year) %>%
  .[[3]] %>%
  sum()


```

Calculating the present value of uneven payments.

```{r pv unequal pv, warning=FALSE,message=FALSE}

cf <- tribble(~Year,~cf,
              1,100,
              2,200,
              3,300,
              4,400,
              5,500)
cf %>%
  mutate(pv = cf/(1+r)^Year) %>%
  .[[3]] %>%
  sum()

```

Calculating the PV with negative cash flow

```{r pv neg cf, warning=FALSE,message=FALSE}

cf <- tribble(~Year,~cf,
              0,-250,
              1,100,
              2,100,
              3,100,
              4,100,
              5,100)
              
cf %>%
  mutate(pv = cf/(1+r)^Year) %>%
  .[[3]] %>%
  sum()


```

Calculating the present value of a finite annuity

```{r finite annuity, warning=FALSE,message=FALSE}

pmt <- 1000
n <- 5
r <- 0.12

cf <- tibble(period = 1:5,
             pmt = pmt)

cf %>%
  mutate(pv = pmt/(1+r)^period) %>%
  .[[3]] %>%
  sum()

```

Calculating the present value infinite annuity

```{r infinite annuity, warning=FALSE,message=FALSE}

pmt <- 1000
r <- 0.12

annuity <- pmt/r
annuity

```

Calculating present value of finite growing annuity

```{r grow annuity,warning=FALSE,message=FALSE}

payment <- 1000
growth <- 0.06
n <- 5
rate <- 0.12

cf <- tibble(period = 1:5,
             pmt = 1000)

cf %>%
  mutate(pmt = payment*(1+growth)^(period - 1)) %>%
  mutate(pv = pmt/(1+rate)^period)

cf %>%
  mutate(pmt = payment*(1+growth)^(period - 1)) %>%
  mutate(pv = pmt/(1+rate)^period) %>%
  .[[3]] %>%
  sum()

```

Calculating present value of infinite growing annuity

```{r grow infinite annuity,warning=FALSE,message=FALSE}

payment <- 1000
growth <- 0.06
rate <- 0.12

annuity <- payment/(rate - growth)
annuity

```

Calculating the Internal rate of return

```{r IRR, warning=FALSE,message=FALSE}

cf <- tibble(Year = 0:5,
             cf = c(-800,200,250,300,350,400))

#using the IRR formula from the jrvFinance library

irr(cf$cf)

```
  
Calculating the loan payments

```{r loan pmt, warning=FALSE,message=FALSE}

car_loan <- 10000
interest <- 0.07
years <- 6

car_payments <- pmt(pv = -car_loan,r = interest, n = years, fv = 0)
car_payments

```

```{r loan pmt schedule, warning=FALSE,message=FALSE}

car_loan <- 10000
interest <- 0.07
years <- 6
car_payments <- pmt(pv = -car_loan,r = interest, n = years, fv = 0)

loan_table <- matrix(ncol = 6, nrow = 6)
loan_table[1,1] <- 1
loan_table[1,2] <- car_loan
loan_table[1,3] <- car_payments
loan_table[1,4] <- car_loan * interest
loan_table[1,5] <- car_payments - (car_loan * interest)
loan_table[1,6] <- car_loan - (car_payments - (car_loan * interest))

for(i in 1:5) {
  
  loan_table[c(i+1),1] <- i + 1
  loan_table[c(i+1),2] <- loan_table[i, 6]
  loan_table[c(i+1),3] <- car_payments
  loan_table[c(i+1),4] <- loan_table[c(i+1),2] * interest
  loan_table[c(i+1),5] <- car_payments - (loan_table[c(i+1),2] * interest)
  loan_table[c(i+1),6] <- loan_table[c(i+1),2] - (car_payments - (loan_table[c(i+1),2] * interest))
  
}

loan_table <- tk_tbl(as.data.frame(loan_table))
colnames(loan_table) <- paste(c("Year", 'Initial_Balance', "Payments", "Interest",
                                "Principal", "Ending_Balance"))

loan_table <- loan_table %>%
  mutate(Ending_Balance = round(Ending_Balance,3))

loan_table
```

Calculating the Future value of deposit in an account

```{r fv value, warning=FALSE,message=FALSE}

#Simple future value

interest <- 0.1
deposit <- 1000

fv_table <- tibble(Year = 1:10,
                   beg_val = 0,
                   int = 0,
                   end_val = 0)

fv_table <- fv_table %>%
  mutate(beg_val = if_else(Year == 1, deposit,beg_val)) %>%
  mutate(int = if_else(Year == 1, beg_val * interest, int)) %>%
  mutate(end_val = if_else(Year == 1, beg_val + int, end_val))

for(i in 2:10) {
  
  fv_table[i,1] <- i
  fv_table[i,2] <- fv_table[c(i-1),4]
  fv_table[i,3] <- fv_table[i,2] * interest
  fv_table[i,4] <- fv_table[i,2] + fv_table[i,3]
  
}

fv_table
  

```

Calculating the future value of regular annual deposits

```{r reg deposit, warning=FALSE,message=FALSE}

interest <- 0.1
annual_deposit <- 1000
n = 10  

fv_table <- tibble(Year = 1:10,
                   beg_value = 0,
                   deposit = 0,
                   int = 0,
                   end_value = 0)

fv_table[1,2] <- 0
fv_table[1,3] <- annual_deposit
fv_table[1,4] <- annual_deposit * interest
fv_table[1,5] <- fv_table[1,3] + fv_table[1,4]


for(i in 2:10) {
  
  fv_table[i,2] <- fv_table[c(i-1),5]
  fv_table[i,3] <- annual_deposit
  fv_table[i,4] <- (fv_table[i,2] + annual_deposit) * interest
  fv_table[i,5] <- fv_table[i,2] + fv_table[i,3] + fv_table[i,4]
  
}

fv_table

```

* Retirement problem

Suppose the current age is 55 years and intention to retire is 60 (5 years to retirement). After retirement, plan is to withdraw $30,000 for the next 8 years. Expected returns on investments is 8%. How much should be deposited each year?

```{r retirement, warning=FALSE,message=FALSE}

# Divide the problem into two parts 
#Part 1 Age 55 to 60
#Part 2 Age 60 to 68

interest <- 0.08
n1 = 5 #part 1
n2 = 8 #part 2
pmt_in_retirement <- 30000

#first calculate amount required to support withdrawing $30000/year during retirement

retirment_amount <- pv(pmt = 30000, r = interest, n = n2, fv = 0, type = 1) #type = 1, withdrawing at the beginning period
retirment_amount

#Calculating the payments required to reach $186191 in retirement

payment <- pmt(pv = 0, fv = retirment_amount, r = interest, n = n1, type = 1)

payment

cat("We need", sprintf("$%.f", payment), "in savings each year to withdraw $30000 in retirement")
```

Calculating interest rates at different compounding periods

```{r comp rates, warning=FALSE,message=FALSE}

interest <- 0.05
initial_dep <- 1000
c_period <- 2 # compounding periods
c_int <- interest/c_period #Interest for compounding period
aft_one_year <- initial_dep * (1 + c_int)^c_period
aft_one_year

cont_comp <- 1000 * exp(interest)
cont_comp

#Compounding for different time period table

cmp_tbl <- tibble(period = c(1,2,10,20,50,100,150,300,800),
                  end_val = 0)

cmp_tbl %>%
  mutate(end_val = initial_dep * (1 + interest/period)^period)

cmp_tbl %>%
  mutate(end_val = initial_dep * (1 + interest/period)^period) %>%
  ggplot(aes(x = period, y = end_val)) +
  geom_line()
    

```

Continuous discounting

```{r continuous disc, warning=FALSE,message=FALSE}



interest <- 0.08

cont_tbl <- tibble(Year = 1:5,
                   cash_flow = seq(100,500,100))

cont_tbl %>%
  mutate(pv = cash_flow * exp(-interest * Year))
  
  
  
```

Calculating Returns from price data

```{r calc returns, warning=FALSE,message=FALSE}

init_dep <- 1000
end_val <- 1200
cmp_period <- 2

#Continuous return
log(end_val/init_dep)

# Implied annual return
((end_val/init_dep)^(1/cmp_period) - 1) * cmp_period

```















